% =====================================================================
%  Section 01 — Layla's Story (Act I) and the Problem Statement
% =====================================================================
\makesectioncover{Layla's Cairo}

\section{Layla's Story}
\label{sec:layla}

\subsection{6:15 AM, Imbaba}

It is six fifteen in the morning. The neighbourhood is already loud.

Layla — \ar{ليلى} — lives in Imbaba on the west bank of the Nile, in
one of the densest neighbourhoods on Earth: \stat{63{,}000 people per
square kilometre}. To put that in perspective, the average density of
Manhattan is about 27{,}000. New Cairo, the satellite city forty
kilometres east, has fewer than 3{,}000 — a \stat{21$\times$ ratio}
between where Cairenes already live and where the state has been
building.

No metro line ran through Imbaba until 2024. None runs at a useful
walking distance from Layla's apartment today.

\begin{callout}
\textbf{\color{plotblue}A note on Layla.}\quad Layla is a \emph{composite
persona} — a fictional commuter who stands in for the
\stat{1.4 million Cairenes} whose commutes the state's formal
transport map does not describe. Every number in her story is real and
notebook-verified: Imbaba's density figure is from \texttt{citypopulation.de}
(Phase 2 source S6, 408 admin rows filtered to Greater Cairo); the
\stat{83-minute} round-trip averages \texttt{explorecity.life}
crowdsourced data and the World Bank Cairo Traffic Congestion Study;
the route polyline is fourteen hand-digitised waypoints across 26th of
July Street and the Corniche.
\end{callout}

\subsection{The office is in Maadi, eleven kilometres away}

Across the city. South of the Nile. \ar{المعادي · ١١ كيلومتر} — eleven
kilometres on paper. The haversine crow-flies distance from Imbaba's
centroid (31.2083\textdegree, 30.0866\textdegree) to her office in
Maadi (31.2581\textdegree, 29.9601\textdegree) is exactly that — eleven
kilometres, computed great-circle with Earth radius \mbox{6{,}371 km}.

But Layla cannot fly. To get to work she must cross two transport
systems that were never designed to connect:

\begin{itemize}
  \item \kw{System A} — the formal, state-built network. Cairo Metro,
        Transport-for-Cairo's published bus routes, the new LRT to the
        New Administrative Capital, the Ring-Road BRT.
  \item \kw{System B} — the lived, informal network. A web of 1{,}302
        microbus stops, 280 \emph{mawaqif} (terminals), and the
        rider-knowledge that every Cairene carries in their head.
\end{itemize}

\subsection{The real route is 83 minutes each way}

\ar{الرحلة الحقيقية · ٨٣ دقيقة}. Three vehicles. Two transfers. No single
app that can plan the trip end-to-end.

\begin{table}[H]
\centering\small
\renewcommand{\arraystretch}{1.25}
\begin{tabular}{@{} >{\color{plotorange}\bfseries}l
                    >{\color{bodytext}}l
                    >{\color{plotyellow}}r @{}}
\toprule
\color{plotblue}Leg & \color{plotblue}Mode & \color{plotblue}Minutes \\
\midrule
Walk to mawfaq         & on foot                  & 8  \\
Wait at mawfaq         & informal terminal queue  & 8.5 \\
Microbus 1             & informal microbus        & 25 \\
Walk between systems   & on foot                  & 12 \\
Metro Line 2 to Maadi  & formal metro             & 30 \\
\midrule
\textbf{Total one-way} &                          & \stat{83} \\
\bottomrule
\end{tabular}
\caption{One morning leg of Layla's commute. Round-trip $\approx$ 166 minutes.
At 260 commuting days per year, Layla loses \stat{720 hours} — about
\stat{30 working days} — to a city that cannot route her.}
\end{table}

\subsection{At 6:15 the mawfaq decides for her}

\ar{الموقف · الساعة ٦:١٥}. The \emph{mawfaq} (Arabic plural \emph{mawaqif})
is the physical expression of System B. Five microbuses are loading at
once, five destinations written in Arabic on taped cardboard, drivers
shouting competing routes.

\begin{itemize}
  \item There is no schedule.
  \item There is no app.
  \item No driver knows her specific trip.
\end{itemize}

She picks a microbus by guessing which direction it starts in. On Phase
1's integrated dataset (cleaning notebook, 280 terminals + Stage-4
SBERT multilingual reconciliation) we can confirm that this specific
mawfaq is not linked to any post-2014 metro station within its 2 km
walkshed. The infrastructure investment of the last decade simply did
not reach her morning.

\subsection{The problem this report is documenting}

\begin{callout}
Cairo runs on \kw{two parallel transport systems}. Since 2014 the state
has spent more than \stat{\$10 billion} expanding System A — Metro Line 3
(Ataba to Adly Mansour, then westwards to Kit Kat), the Cairo LRT to the
New Administrative Capital, and the Ring-Road BRT. Yet the daily reality
of the \stat{1.4 million Cairenes who commute like Layla} has not
materially changed.

Why? Because System A was largely built where people \emph{will} live
according to state plans, while System B continues to carry people
where they \emph{already} live. The two systems overlap geographically
but not functionally — neither has a published timetable that
references the other; neither can route the rider through the other.

This report investigates that disconnect with data. \kw{Phase 1} cleans
seven raw GeoJSON datasets that describe the \emph{existing} transport
reality (1{,}302 stops, 1{,}784 routes, 280 terminals, 9{,}258 road
segments, 1{,}525 population hexagons), answers twelve research
questions about it, and produces \kw{four structural gaps} —
115 ghost terminals, 19 empty-return terminals, vehicle-route mismatch
on 75\% of long routes, and 79 underserved population hexagons.

\kw{Phase 2} then scrapes eight new sources to describe what the state
\emph{has built} since 2014 (89 metro stations, 20 LRT stations,
21 BRT stations, 68 districts of Greater Cairo population, vehicle
ownership, GTFS, OpenStreetMap), integrates it with the Phase 1 reality
through a four-stage pipeline (KNN spatial $\to$ OSM cross-verification
$\to$ RapidFuzz fuzzy text $\to$ \kw{SBERT multilingual semantic
matching}), then asks thirteen research questions and tests
\kw{three formal hypotheses} about whether the new infrastructure
closes the four gaps.

The answer — derived from a \stat{\$10B} investment cross-tabulated against
four gaps and three modes — is that \stat{no cell of the resulting
3$\times$4 matrix sits above 25\%}. Most are below 16\%. The
infrastructure did not reach Layla. \kw{The data shows why, and the
report walks through every step of how we know.}
\end{callout}

\subsection{How this report is organised}

Reading order:

\begin{enumerate}
  \item \textbf{Phase 1} — what Cairo's existing transport system looks like.
        Data sources, pre-clean nulls, KNN imputation, seven spatial
        merges, twelve research questions, four structural gaps.
        (Sections 2--8.)
  \item \textbf{Phase 2} — what the state built since 2014. Eight scraped
        sources, the four-stage integration pipeline, thirteen
        research questions. (Sections 9--13.)
  \item \textbf{AI techniques} — three explicit AI methods used in the
        project (SBERT, K-Means, Moran's~I), each chosen because a
        classical method could not solve the same problem. (Section 14.)
  \item \textbf{Hypothesis tests} — a standalone methods reference for
        every statistical test used (Spearman, Kruskal-Wallis,
        Mann-Whitney, Cliff's $\delta$, $\varepsilon^2$, permutation,
        Moran's~I, K-Means), then a deep walkthrough of \kw{H1}
        coverage-need mismatch, \kw{H2} LRT catchment deficit, and
        \kw{H3} BRT corridor match. (Section 15.)
  \item \textbf{Synthesis} — what the evidence collectively says, the
        gap-closure matrix that is the single most important result,
        and the resolution stat (\stat{83 $\to$ 40 minutes}) that ends
        Layla's story. (Section 16.)
  \item \textbf{Appendices} — file manifest and a glossary of every
        abbreviation. (Sections 17--18.)
\end{enumerate}
