% =====================================================================
%  Section 02 — Phase 1 introduction (verbatim user paraphrase)
% =====================================================================
\makesectioncover{Phase 1 \textbullet\ The Existing Reality}

\section{Phase 1 \textbullet\ The Existing Reality}
\label{sec:phase1-intro}

In this phase we are trying to know what the existing transport is.
It is an auditing phase before discussing anything regarding the new
infrastructure and whether they are helpful or not.

The work is divided into two notebooks:

\begin{itemize}
  \item \fn{Cleaning Data.ipynb}: here we simply load seven raw
        GeoJSON datasets, we then reproject them into a common CRS,
        we use KNN to impute missing values, and we build seven
        spatial attribute merges that will be used further on in the
        analysis section.
  \item \fn{Visual Analysis.ipynb}: here we are answering twelve
        exploratory questions by using the merges from the first
        cleaning notebook, and we produce 40 Plotly visualisations.
        The questions result in 4 gaps that are the base for Phase 2
        hypotheses; the 4 gaps are ghost terminals, empty-return
        terminals, vehicle-route mismatch, and underserved hexagons.
\end{itemize}

\subsection{Two-system framing}

In this report we will use two important labels:

\begin{itemize}
  \item \kw{System A} (formal), which is the published network. In
        Phase 1 we approximate it by routes with fares and the
        published \fn{Public Transport Terminals.json}. However, in
        Phase 2 this is mixed with GTFS, the metro network website,
        the LRT and the BRT.
  \item \kw{System B} (informal), this is what the city actually uses
        every morning. In Phase 1 we brought its data from
        \fn{Cairo Daily Boarding.json} and the passenger-flow segments
        \fn{Public Transport Passenger Flow.json}.
\end{itemize}

The whole idea behind Phase 1 is to ensure the two systems can be
compared on the same map under the same set of metrics, such as
distances, boarding totals, and fare per km. The 4 structural gaps are
the cells where the comparison fails.
