% =====================================================================
%  Section 17 — Appendix: file manifest
% =====================================================================
\appendix
\makesectioncover{Appendix · File Manifest}

\section{File Manifest}
\label{sec:appendix-files}

The complete list of files produced and consumed by the project.
Reproducibility = every number in the report can be traced to one of
these files.

\subsection{Phase 1 raw inputs}

\fn{DatasetsGeojson/}
\begin{itemize}\setlength{\itemsep}{0pt}
  \item \fn{Cairo Daily Boarding.json} (1{,}302 rows, Point)
  \item \fn{Public Transport Routes.json} (1{,}784 rows, LineString)
  \item \fn{Public Transport Terminals.json} (280 rows, Point)
  \item \fn{Public Transport Passenger Flow.json} (9{,}258 rows, LineString)
  \item \fn{Public Transport Vehicles Flow.json} (5{,}592 rows, LineString)
  \item \fn{Public Transport Commercial Speeds 2.json} (26{,}154 rows, LineString)
  \item \fn{Population \& Employment Access.json} (1{,}525 rows, Polygon H3 hex)
\end{itemize}

\subsection{Phase 1 cleaned outputs}

\fn{CleanedData/} — 21 CSV files, $\sim$41 MB total.

\textbf{Cleaned individual datasets:}
\fn{cleaned\_boarding.csv}, \fn{cleaned\_population.csv},
\fn{cleaned\_routes.csv}, \fn{cleaned\_speeds.csv},
\fn{cleaned\_terminals.csv}, \fn{cleaned\_passenger\_flow.csv},
\fn{cleaned\_vehicle\_flow.csv}.

\textbf{Merged tables:}
\fn{merged\_A\_boarding\_population.csv},
\fn{merged\_B\_terminals\_population.csv},
\fn{merged\_D\_terminal\_flow\_ratios.csv},
\fn{merged\_E\_terminal\_pop\_flow.csv},
\fn{merged\_F\_routes\_per\_terminal.csv},
\fn{merged\_G\_underused\_terminals.csv}.

\textbf{Question-specific helpers:}
\fn{q3\_drop\_points.csv}, \fn{q4\_formal\_vs\_informal.csv},
\fn{q5\_hex\_terminal\_counts.csv}, \fn{q6\_routes\_vs\_passengers.csv},
\fn{q7\_morning\_vs\_evening.csv},
\fn{q8\_spatial\_autocorrelation.csv}, \fn{q9\_underserved.csv},
\fn{q12\_vehicle\_type\_route\_length.csv}.

\subsection{Phase 1 exports}

\fn{Exports/} — 40 standalone Plotly HTML files (Q1a, Q1b, Q2a, \ldots,
Q12e). Every one has a static PNG render in
\fn{Report/figures/phase1/}.

\subsection{Phase 2 raw inputs}

\fn{Phase2/RawData/}
\begin{itemize}\setlength{\itemsep}{0pt}
  \item \fn{gtfs\_bus\_metro.zip} + \fn{gtfs\_bus\_metro/} (S1)
  \item \fn{wiki\_metro\_masterlist.html} (S3)
  \item \fn{wiki\_lrt.html} (S4)
  \item \fn{gmaps\_brt\_brt\_en.html} + \fn{gmaps\_brt\_trd\_ar.html} (S5)
  \item \fn{citypop\_egypt\_admin.html} + \fn{nominatim\_districts.json} (S6)
  \item \fn{wiki\_vehicles\_per\_capita.html} +
        \fn{wb\_motor\_vehicles\_per\_1k.json} +
        \fn{wb\_passenger\_cars\_per\_1k.json} (S7)
  \item \fn{osm\_overpass.json} + \fn{osm\_lrt\_per\_station.json} (S8)
\end{itemize}

\subsection{Phase 2 cleaned outputs}

\fn{Phase2/CleanedData/}
\begin{itemize}\setlength{\itemsep}{0pt}
  \item GTFS: \fn{gtfs\_stops.csv}, \fn{gtfs\_routes.csv},
        \fn{gtfs\_trips.csv}, \fn{gtfs\_shapes.csv},
        \fn{gtfs\_fare\_attributes.csv}.
  \item \fn{metro\_stations.csv} (89 rows, S3).
  \item \fn{lrt\_stations.csv} (20 rows, S4 with backfill).
  \item \fn{brt\_stations.csv} (21 rows, S5 deduplicated).
  \item \fn{districts.csv} (long), \fn{districts\_wide.csv} (S6).
  \item \fn{vehicles\_by\_governorate.csv} (S7).
  \item \fn{osm\_features.geojson} (S8).
  \item \fn{null\_audit\_before.csv}, \fn{null\_audit\_after.csv}.
  \item \fn{q17\_dense\_underserved\_hexes.csv}.
\end{itemize}

\subsection{Phase 2 integrated}

\fn{Phase2/Integrated/}
\begin{itemize}\setlength{\itemsep}{0pt}
  \item \fn{matched\_pairs.csv} — every Phase 1 $\to$ Phase 2 station
        pair with stage of origin, similarity, discrepancy distance.
  \item \fn{Phase2\_integrated.geoparquet} — the merged spatial layer.
\end{itemize}

\subsection{Phase 2 exports}

\fn{Phase2/Exports/} — 117+ standalone HTML files across three layers:

\begin{itemize}
  \item Root level (45 files) — Q13--Q25, H1--H3, hero-map components,
        market-sizing visuals.
  \item \fn{Phase2/Exports/notebook\_visuals/} (71 files) — granular
        per-chart exports.
  \item \fn{Phase2/Exports/notebook\_sections/} (13 files) — bundled
        per-question aggregations.
\end{itemize}

\subsection{Phase 2 hero maps}

Three standalone Folium maps at \fn{Phase2/} root:

\begin{itemize}
  \item \fn{two\_cairos\_map.html} — formal vs informal overlay.
  \item \fn{headline\_coverage\_need\_map.html} — coverage-need
        synthesis.
  \item \fn{adly\_mansour\_zoom.html} — convergence-node case study.
\end{itemize}

\subsection{Report assets}

\fn{Report/}
\begin{itemize}\setlength{\itemsep}{0pt}
  \item \fn{main.tex}, \fn{preamble.tex}, \fn{build.sh},
        \fn{README.md}.
  \item \fn{sections/00\_titlepage.tex} \ldots
        \fn{sections/18\_appendix\_glossary.tex} (19 .tex files).
  \item \fn{figures/phase1/} (37 PNGs) +
        \fn{figures/phase2/} (45 PNGs) +
        \fn{figures/phase2/notebook/} (48 PNGs) +
        \fn{figures/hero/} (3 PNGs) = \stat{133 figures total}.
  \item \fn{scripts/htmls\_to\_pngs.py} — selenium-based universal
        HTML$\to$PNG renderer.
\end{itemize}

\subsection{Story-app source (referenced for the Layla narrative)}

\fn{cairo\_story\_app/}
\begin{itemize}\setlength{\itemsep}{0pt}
  \item \fn{data/findings.py} — single source of truth for every
        statistic the report cites.
  \item \fn{chapters/chapters\_config.py} — 14-chapter narrative arc;
        Layla story chapters 0--3 + 7 are quoted in
        \S\ref{sec:layla}.
  \item \fn{findings.md} — markdown copy of \fn{findings.py}.
\end{itemize}
